---
title: "Data Cleaning - GIS Final Project"
author: "Ben Steves"
date: "4/26/2021"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(mosaic)
require(tidyverse)
```

This is all of the data transforming that I did, mostly prior to doing any maps. I added in the vehicles dataset after I finished my IDW and LPI methods, though. 


### Read in data

Route data

```{r}
routes <- read.csv("https://docs.google.com/spreadsheets/d/e/2PACX-1vRcpyEUb1ZT5UaGkvlIeRpjgzebye5ZYN4yFfuzxUUJiif5z8LD9UlEH0REXyJkctL9rDTSr8o4BVlG/pub?output=csv")
```

State observation data

```{r}
michigan <- read.csv("Michiga.csv")
illinois <- read.csv("Illinoi.csv")
indiana <- read.csv("Indiana.csv")
ohio <- read.csv("Ohio.csv")
wisconsin <- read.csv("Wiscons.csv")
```

Vehicle noise and car total data

```{r}
vehicles <- read.csv("VehicleData.csv")
```

```{r}
#migrants <- read.csv("MigrantSummary.csv")
```


### Renaming variables

States were coded into the dataset as numbers, so I made new variable for state name that wasn't a number.

```{r}
routes <- routes %>%
 # mutate(StateNum == as.character(StateNum)) %>%
  mutate(StateName = case_when(
    StateNum == 49 ~ "MI",
    StateNum == 34 ~ "IL",
    StateNum == 35 ~ "IN",
    StateNum == 66 ~ "OH",
    StateNum == 91 ~ "WI"
  )) %>%
  filter(StateName %in% c("IL", "IN", "MI", "OH", "WI"))
```

### Put all states data into one mega-dataset

```{r}
midwest <- ohio %>%
  bind_rows(wisconsin) %>%
  bind_rows(michigan) %>%
  bind_rows(illinois) %>%
  bind_rows(indiana)
```

### Looking to see which birds have most observations

Only printed first 6 values from data

```{r}
midwest %>%
  filter(AOU %in% 1000:2000) %>%
  select(AOU, SpeciesTotal) %>%
  group_by(AOU) %>%
  summarize(n = sum(SpeciesTotal)) %>%
  arrange(desc(n)) %>% head(6)
```

The dataset had tons of birds, but all of the waterfowl had AOU species numbers from 1000-2000, which were values that I needed. Metadata shows all of the species and their AOU values, but from the birds I'm looking at:

Canada goose - 1720
Wood duck - 1440
Mute swan - 1782
Mallard - 1320
Blue winged teal - 1400
Common Merganser - 1290

```{r}
# michigan %>%
#   filter(AOU == 1290) %>%
#   filter(Year == 2000) %>%
#   select(AOU, SpeciesTotal) %>%
#   summarize(n = sum(SpeciesTotal))
```

### Creating new index variables for birds, and adding new variable for actual name. 

```{r}
midwest <- midwest %>%
  filter(AOU %in% c(1320, 1290, 1400, 1440, 1720, 1782)) %>%  
  mutate(AOU_idx = case_when(
    AOU == 1320 ~ 1,
    AOU == 1290 ~ 2,
    AOU == 1400 ~ 3,
    AOU == 1440 ~ 4,
    AOU == 1720 ~ 5,
    AOU == 1782 ~ 6
  )) %>%
  mutate(SpeciesName = case_when(
    AOU_idx == 1 ~ "Mallard",
    AOU_idx == 2 ~ "Common Merganser",
    AOU_idx == 3 ~ "Blue Winged Teal",
    AOU_idx == 4 ~ "Wood Duck",
    AOU_idx == 5 ~ "Canada Goose",
    AOU_idx == 6 ~ "Mute Swan"
  ))
```

### Changing states to text in midwest dataset as well, so I can easily merge it with routes.csv

```{r}
midwest <- midwest %>%
  #mutate(StateNumChar == as.character(StateNum)) %>%
  mutate(StateName = case_when(
    StateNum == 49 ~ "MI",
    StateNum == 34 ~ "IL",
    StateNum == 35 ~ "IN",
    StateNum == 66 ~ "OH",
    StateNum == 91 ~ "WI"
  )) %>%
  filter(StateName %in% c("IL", "IN", "MI", "OH", "WI"))
```

### Filtering only years from 2009-2019 and summing the number of birds by species

This is what I originally did for my IDW

```{r}
midwest_2009_2019 <- midwest %>%
  filter(Year %in% 2009:2019) %>%
  group_by(Route, StateName, SpeciesName, AOU_idx) %>%
  summarize(n = sum(SpeciesTotal))
```

### Filtering only years from 2009-2019 and averaging the number of birds by species

Improved for my LPI, using means number of birds at a route for all years 2009-2019. 

```{r}
means_midwest_2009_2019 <- midwest %>%
  filter(Year %in% 2009:2019) %>%
  group_by(Route, StateName, SpeciesName, AOU_idx) %>%
  summarize(MeanBirds = mean(SpeciesTotal))
```

### Joining datasets

Joining the states datasets with the routes datasets. Links the bird observation data from the states datasets with the geospatial elements of routes dataset. 

```{r}
mean_routes_joined <- means_midwest_2009_2019 %>%
    left_join(routes, by = c("Route", "StateName"))
```

```{r}
routes_joined <- midwest_2009_2019 %>%
  left_join(routes, by = c("Route", "StateName"))
```

### Working in vehicles dataset. 

Exporting it to do some extra cleaning in excel, where I calculate the total cars seen per route and the number of routes that had excessive noise. Filtered only for years 2009-2019 as well. 

```{r}
vehicles <- vehicles %>%
  filter(StateNum %in% c(49, 34, 35, 66, 91)) %>%
  filter(Year %in% 2009:2019)
```

```{r}
write.csv(vehicles, file = "vehicles_nosum.csv")
```

```{r}
vehicles_sums <- read.csv("vehicles_consumz.csv")
```

###Changing states to text in vehicles dataset as well

```{r}
vehicles_sums <- vehicles_sums %>%
  #mutate(StateNumChar == as.character(StateNum)) %>%
  mutate(StateName = case_when(
    StateNum == 49 ~ "MI",
    StateNum == 34 ~ "IL",
    StateNum == 35 ~ "IN",
    StateNum == 66 ~ "OH",
    StateNum == 91 ~ "WI"
  ))
```

### Calculate mean total number of cars, averaged across years 2009-2019

```{r}
vehicles_means <- vehicles_sums %>%
 group_by(Route, StateName) %>%
  summarize(CarMeanTotal = mean(CarTotal))
```

### Calculate mean noise levels, averaged across years 2009-2019

```{r}
noise_means <- vehicles_sums %>%
  group_by(Route, StateName) %>%
  summarize(NoiseMean = mean(NoiseTotal))
```

### Joining noise and car data

```{r}
means_joined <- noise_means %>%
  left_join(vehicles_means, by = c("Route", "StateName"))
```

### Joining vehicle and route/observation data

```{r}
fullData <- mean_routes_joined %>%
  left_join(means_joined, by = c("Route", "StateName"))
```

### Exporting data for use in GWR

```{r}
write.csv(fullData, file = "gwrdata.csv")
```






